Global code motion of congruent instructions

### 2017-04-18 paper.tex
tweak some algorithms to stand uncut on a page

// Return true if the merge-set of source node of a visited J-edge (incoming edge of
// lnode) is not the subset of the merge-set of lnode.
bool still_inconsistent(Node lnode, JEdges JE, Visited V, MergeSet Merge)
  for e in (all incoming edges to lnode)
    if (e is in JE && e in V)
      if (Merge(Source Node of e) is not a subset of Merge(lnode))
        return true
  return false

// Compute a merge-set top-down in breadth first order.
bool constructMergeSet_1(BFSList B, JEdges JE, DomLevel DL)
  repeat = false
  V = empty list of visited edges
  for n in B
    for e in (all incoming edges to n)
      if (e is in JE && e not in V)
        V(e) = true
        let tmp = Source Node of e
        let tnode = Target Node of e
        let lnode = NULL
        while (DL(tmp) >= DL(tnode))
          Merge(tmp) = Merge(tmp) U Merge(tnode) U {tnode}
          lnode = tmp
          tmp = dom-parent(tmp)
        repeat = still_inconsistent(lnode, JE, Merge)
  return repeat

@@

  Ms(N) = null
  // Mergeset of N is the union of merge sets of its successors
  for W in successors(N)
    Ms(N) = Ms(N) U Mr(W)

  // Iterate over all the uses of A and see if any intersect with the merge-set of N.
  for T in users(A)
    B = basic_block(T)
    while (B != null and B != def_bb(A))
      if B in Ms(N) return true
      B = dom-parent(B)
  return false

@@

\begin{verbatim}
computing downward-safety of hoistable instructions at hoist-point
  Worklist = list of all basic blocks of hoistable instructions
  for each basic block B in the dominator tree starting at hoist-point:
    if Worklist is empty
      return false // Path exists! not downward safe
    if B is in Worklist:
      remove B from worklist
      remove subtree with root at B // Available at B => downward safe from B
    if B is a leaf basic block:
      return false  // Path exists! not downward safe
    if B dominates hoist-point:
      return false  // Back edge
  return true
\end{verbatim}

@@

\begin{verbatim}
void doGCM()
  Analyses available: Dominator Tree, DFS Numbering, Memory SSA,
  Compute DJ-graph of function and compute mergeset based on that
  do
    Compute GVN of each expression in the function
    int N = doCodeHoisting()
  while (N > 0) // If any of candidates were hoisted then repeat hoisting
  doCodeSinking()
\end{verbatim}

@@

% Explain the overall algorithm, explain how anticipability is computed on demand
\begin{verbatim}
int doCodeHoisting()
  For each value number VN with 2 or more instructions:
    sort the instructions according to DFSIn numbers
    if first two I1, I2 are hoistable:
      if they are downward-safe at hoist-point:
        if they are profitable to be hoisted at hoist-point:
          proceed to check ability to hoist of subsequent instruction with same VN
        else
          proceed to check if I2 and I3 (if available) can be hoisted

  For each set of instructions (I1, ..., Ik) with same VN that can be hoisted:
    move the first one I1 to the hoist-point
    update the use of all other candidates (I2, ..., Ik) to refer to I1
    remove all other instructions (I2, ..., Ik)
    if I1 has memory references (a load, store etc.)
      update MemorySSA of I1 and others to point to the MemorySSA reference of I1
      remove MemorySSA reference of all others which were deleted
    update statistics

  return number_of_hoisted_candidates
\end{verbatim}

@@

\begin{verbatim}
void doCodeSinking()
  For each value number VN with 2 or more instructions:
    if there are two instructions with same immediate successor as post-dominator
      if both are only used in the same PHI-node (PN) of the successor
        if both have dependencies that are available or can be made available at the sink point
           mark VN as sinkable

  For each pair I1, I2 of sinkable instructions:
    Move the first one I1 to the sink point which is just after all the PHI-nodes
    in the post-dominator
    update all the uses of PHI (PN) with I1
    Remove I2 and PN
    if I1 is a memory reference:
      update MemorySSA of I1 and others to point to the MemorySSA reference of I1
      remove MemorySSA reference of all others which were deleted
    update statistics

>>>>

// Return true if the merge-set of source node of a visited J-edge (incoming edge of
// lnode) is not the subset of the merge-set of lnode.
bool still_inconsistent(Node lnode, JEdges JE, Visited V, MergeSet MergeSet)
  for e in (all incoming edges to lnode)
    if (e is in JE && e in V)
      if (MergeSet(Source Node of e) is not a subset of MergeSet(lnode))
        return true
  return false

// Compute a merge-set top-down in breadth first order.
bool constructMergeSet_1(BFSList B, JEdges JE, DomLevel DL)
  repeat = false
  V = empty list of visited edges
  for n in B
    for e in (all incoming edges to n)
      if (e is in JE && e not in V)
        V(e) = true
        let tmp = Source Node of e
        let tnode = Target Node of e
        let lnode = NULL
        while (DL(tmp) >= DL(tnode))
          MergeSet(tmp) = MergeSet(tmp) U MergeSet(tnode) U {tnode}
          lnode = tmp
          tmp = dom-parent(tmp)
        repeat = still_inconsistent(lnode, JE, MergeSet)
  return repeat

@@

  MergeSet(N) = null
  // The merge-set of N is the union of merge-sets of its successors.
  for W in successors(N)
    MergeSet(N) = MergeSet(N) U MergeSet(W)

  // Iterate over all the uses of A and see if any intersect with the merge-set of N.
  for T in users(A)
    B = basic_block(T)
    while (B != null and B != def_bb(A))
      if B in MergeSet(N) return true
      B = dom-parent(B)
  return false

@@

\begin{verbatim}
computing downward-safety of hoistable instructions at hoist-point
  Worklist = list of all basic blocks of hoistable instructions
  for each basic block B in the dominator tree starting at hoist-point:
    if Worklist is empty, return false // Path exists! not downward safe
    if B is in Worklist:
      remove B from worklist
      remove subtree with root at B // Available at B => downward safe from B
    if B is a leaf basic block, return false  // Path exists! not downward safe
    if B dominates hoist-point, return false  // Back edge
  return true
\end{verbatim}

@@

\begin{verbatim}
void doGCM()
  Analyses available: Dominator Tree, DFS Numbering, Memory SSA,
  Compute DJ-graph of function and compute mergeset based on that
  do
    Compute GVN of each expression in the function
  while (doCodeHoisting() > 0) // If any of candidates were hoisted then repeat hoisting
  doCodeSinking()
\end{verbatim}

@@

\begin{verbatim}
int doCodeHoisting()
  For each value number VN with 2 or more instructions (I1, I2, ..., Ik)
    sort the instructions according to DFSIn numbers
    if first two I1, I2 are hoistable and if they are downward-safe at hoist-point
       and if they are profitable to be hoisted at hoist-point
      proceed to check ability to hoist of subsequent instruction with same VN
    else
      proceed to check if I2 and I3 (if available) can be hoisted
  For each set of instructions (I1, ..., Ik) with same VN that can be hoisted:
    move the first one I1 to the hoist-point
    update the use of all other candidates (I2, ..., Ik) to refer to I1
    remove all other instructions (I2, ..., Ik)
    if I1 has memory references (a load, store, call, etc.)
      update MemorySSA of I1 and others to point to the MemorySSA reference of I1
      remove MemorySSA reference of all others which were deleted
    update statistics
  return number of hoisted instructions
\end{verbatim}

@@

\begin{verbatim}
void doCodeSinking()
  For each value number VN with 2 or more instructions
    if there are two instructions with same immediate successor as post-dominator
      if both are only used in the same PHI-node (PN) of the successor
        if both have dependencies that are available or can be made available at the sink point
           mark VN as sinkable
  For each pair (I1, I2) of sinkable instructions
    Move I1 to the sink point which is just after all the PHI-nodes in the post-dominator
    update all the uses of PHI node (PN) with I1
    Remove I2 and PN
    if I1 is a memory reference:
      update MemorySSA of I1 and others to point to the MemorySSA reference of I1
      remove MemorySSA reference of all others which were deleted
    update statistics
\end{verbatim}
